HUME: Human UCCA-based Machine-Translation Evaluation

### 2016-03-15 paper/main.tex
new iaa, time

\subsection{Annotation Statistics}

For the annotation of MT output, we recruited seven annotators, two for each of the target languages apart from Czech. They were
all native speakers of the target languages and fluent in English. For Czech, we also recruited a  group of additional annotators
to annotate small numbers of sentences each for the purpose of calculating inter-annotator agreement. In \tabref{tab:annot}
we show the total number of sentences and nodes annotated by each annotator, where we consider the additional Czech
annotators as a single annotator. Note that not all nodes in all sentences were annotated. Sometimes this was because the
annotator did not finish a node, or accidentally missed a node, but in many cases it was also because it was an implicit 
UCCA node, which was not shown to the annotators since there is no corresponding source segment.
\begin{table}
\begin{center}
{\small
\begin{tabular}{ll|cccc}
& & cs & de & pl & ro \\
\hline
Sentences &  Annot. 1 & 324   & 339  & 351  & 230  \\
 & Annot. 2 & -- & 104  & 340  & 337 \\
\hline
Nodes & Annot. 1 & 8794  & 9253 & 9557  & 6152 \\
 &Annot. 2 & -- & 2906  & 9303  & 9228  \\
\end{tabular}
}
\caption{Total numbers of sentences and nodes annotated on the MT output.}
\label{tab:annot}
\end{center}
\end{table}

\emph{Timing of annotations}

@@

\subsection{Inter-Annotator Agreement}
\label{sec:iaa}
\begin{itemize}
  \item Multiple annotation stats
  \item Kappas -- ROG, AB
  \item Kappa versus length
  \item Kappa versus bleu?
\end{itemize}

@@

In order to assess the consistency of the annotation, we measure the Inter-Annotator
Agreement (IAA) using Cohen's Kappa on the multiply-annotated portions of the data.
For German, Polish and Romanian we used two annotators, so we report IAA on the overlap
between these annotators. For Czech we used one main annotator, plus several other extra
annotators to provide IAA.

To calculate Kappa, we consider the annotation task as a classification task on 
nodes. We only consider nodes where there are at least two annotations from different
annotators and we treat nodes with more than two annotations as a set of paired
annotations\bh{This may not be correct, so should/could we do Kappa across the 
variable multiple annotations on Czech}. We consider Kappa separately on the lexical
nodes (annotated as red, orange or green) and the structural nodes (annotated
as acceptable or bad) since there should not be any confusion between these two types of
nodes. The small number of nodes which one annotator had labelled as structural and 
the other as lexical were ignored for the purposes of IAA, since such cases of annotator 
error could be caught by improved tool support.  The Kappas are shown in \tabref{tab:iaa}
\begin{table}[!ht]
\begin{center}
\begin{tabular}{l|cccc}
 & cs & de & pl & ro \\
\hline
Sentences & -- & 102 & 334 & 217 \\
\hline
Lexical nodes & -- & 1724 & 5386 & 3570 \\
Kappa & -- & 0.29 & 0.54 & 0.50 \\
\hline
Structural nodes & & 1040 & 2655 & 1989 \\
Kappa & -- & 0.44 & 0.33 & 0.58 \\
\end{tabular}
\caption{Inter-annotator agreement for the multiply-annotated portions of the data, as
measured by Cohen's Kappa. }
\label{tab:iaa}
\end{center}
\end{table}

To assess whether, as we hoped, HUME would make it easier to reliably annotate long sentences,
we binned the sentences according to length and measured Kappa on each of the bins. Each bin covers
a length range of 10, and we show the variance of kappa with length in \figref{fig:iaalength}.\bh{Note that cs is missing}

@@

\subfloat[English-Czech]{
  \iaafig{de}
}
&
\subfloat[English-German]{
  \iaafig{de}

@@

}
\end{tabular}
\caption{Inter-annotator Agreement (Kappa) versus sentence length, separately for
structural and lexical nodes. }
\label{fig:iaallength}

>>>>

\subsection{Annotation Statistics}
\label{sec:annot_stats}

For the annotation of the MT output, we recruited two annotators for each of German, Romanian
and Polish and one main annotator for Czech. In order to allow us to calculate IAA also for
the Czech annotations we recruited several further annotators to work on a small number of 
sentences each. We treat these further annotators as one annotator, giving us two annotators
for each language pair.
The annotators were all native  speakers of the target languages and fluent in English.

In \tabref{tab:annot}
we show the total number of sentences and nodes annotated by each annotator. Note that not all nodes in all sentences were annotated. Sometimes this was because the
annotator did not finish a node, or accidentally missed a node, but in many cases it was also because it was an implicit 
UCCA node, which was not shown to the annotators since there is no corresponding source segment.
\begin{table}
\begin{center}
{\small
\begin{tabular}{ll|cccc}
& & cs & de & pl & ro \\
\hline
Sentences &  Annot. 1 & 324   & 339  & 351  & 230  \\
 & Annot. 2 & 205 & 104  & 340  & 337 \\
\hline
Nodes & Annot. 1 & 8794  & 9253 & 9557  & 6152 \\
 &Annot. 2 & 5553 & 2906  & 9303  & 9228  \\
\end{tabular}
}
\caption{Total numbers of sentences and nodes annotated on the MT output.}
\label{tab:annot}
\end{center}
\end{table}

The annotators worked remotely, at their own pace, so we did not collect precise timing
information. Nevertheless we are able to estimate the annotation time using the timestamps
provided by the annotation tool, which are recorded whenever an annotated sentence is
submitted. Annotators are not able to re-open a sentence once submitted.

To estimate the annotation time, we calculate the time difference between successive 
sentences, and discard outlying times since we assume there was a break taken. From inspection
of histograms of annotation times, we set the upper threshold at 500 seconds. We
show median annotation times in Table~\ref{tab:annot_times}, indicating that the annotation
of a sentence takes around 2--4 minutes, with some variation between annotators.

\begin{table}
\begin{center}
\begin{tabular}{l|cccc}
& cs & de & pl & ro \\
\hline
Annot. 1 & 255 & 140  & 138 & 96 \\
Annot. 2 & -- & 162 & 229 & 207 \\
\end{tabular}
\caption{Median annotation times per sentence, in seconds.
We do not give times for the second Czech 
``annotator'', since this was a collection of annotators, working in parallel.}
\label{tab:annot_times}
\end{center}
\end{table}

@@

\subsection{Inter-Annotator Agreement}
\label{sec:iaa}

@@

In order to assess the consistency of the annotation, we measure the Inter-Annotator
Agreement (IAA) using Cohen's Kappa on the multiply-annotated portions of the data.
For German, Polish and Romanian we used two annotators, with a significant overlap
between the sentences they annotated. For Czech we had one main annotator, plus
several second annotators, so we treat all these second annotators as a single
annotator.

To calculate Kappa, we consider the annotation task as a classification task on 
nodes, and consider all nodes where there are two annotations from different
annotators We report the overall Kappa, as well as separate Kappas on lexical
nodes (annotated as red, orange or green) and the structural nodes (annotated
as acceptable or bad). In general there should not be confusion between the two
types of nodes, and in future work the separation between the nodes should be enforced
by the tool.  The Kappas are shown in \tabref{tab:iaa}
\begin{table}[!ht]
\begin{center}
\begin{tabular}{l|cccc}
 & cs & de & pl & ro \\
\hline
Sentences & 181 & 102 & 334 & 217 \\
\hline
All nodes & 4686   & 2793   & 8384   & 5604  \\
Kappa & 0.64   & 0.61   & 0.58   & 0.69  \\
\hline
Lexical nodes & 2982 & 1724 & 5386 & 3570 \\
Kappa & 0.54 & 0.29 & 0.54 & 0.50 \\
\hline
Structural nodes & 1602 & 1040 & 2655 & 1989 \\
Kappa & 0.31 & 0.44 & 0.33 & 0.58 \\
\end{tabular}
\caption{Inter-annotator agreement for the multiply-annotated portions of the data, as
measured by Cohen's Kappa. }
\label{tab:iaa}
\end{center}
\end{table}

To assess whether, as we hoped, HUME would make it easier to reliably annotate long sentences,
we binned the sentences according to length and measured Kappa on each of the bins. Each bin covers
a length range of 10, and we show the variance of kappa with length in \figref{fig:iaalength}.

@@

\subfloat[English-Czech]{
  \iaafig{cs}
}
&
\subfloat[English-German]{
  \iaafig{de}

@@

}
\end{tabular}
\caption{Inter-annotator Agreement (Kappa) versus sentence length, separately for
structural and lexical nodes. }
\label{fig:iaalength}
